Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-06-03 paper.Rmd
adjust Table 2 results for N differences

library(readr)
library(dplyr)
library(tidyr)
library(kableExtra)
```

@@

```{r table2-smooth}
table2_smooth <- read_csv("data/table2-smooth.csv")

table2_smooth %>%
  dplyr::select(-Model) %>%
  rename(Horizon = horizon) %>%
  knitr::kable(
    digits = 2, booktabs = TRUE, 
    caption = "Test set AUC-ROC for Models including PITF; original version based on smoothed ROC curves") 
```

@@

table2_nosmooth %>%
  dplyr::select(-Model) %>%
  rename(Horizon = horizon) %>%
  knitr::kable(
    digits = 2, booktabs = TRUE, 
    caption = "Test set AUC-ROC for Models including PITF; version \\textit{without} smoothed ROC curves") 
```

>>>>

library(readr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(stringr)

# function to adjust the "cases" labels in Table 2
fix_case_labels <- function(x) {
  # escape the opening and closing parentheses
  stringr::str_replace_all(x, "([\\(\\)])", "\\\\\\1")
}

# function to adjust the model/column labels in Table 2
```

Issues in the B&S replication:

- The AUC-ROC statistics are calculated using smoothed ROC curves
- The "Weighted by PITF" model is incorrectly implemented
- The "PITF Split Population" model is incorrectly implemented
- The forecasts are scored using incidence, not onset

Pre-registration should have had:

- The escalation model specification, prior to looking at data!
- The hyper-parameter tuning strategy.

With rare events, also a lot of variance over RNG state.

@@

@@

table2_nosmooth %>%
  dplyr::select(-Model, -horizon) %>%
  knitr::kable(
    digits = 2, booktabs = TRUE, 
    caption = "Replication of BS Table 2: Test set AUC-ROC for Escalation model with and without structural PITF contribution") %>%
  column_spec(2:6, width = "2cm") %>%
  pack_rows(index = c("One-month forecasts" = 1, "Six-month forecasts" = 1)) %>%
  footnote("Differences from the original BS Table 2: (1) AUC-ROC values are computed on the common subset of cases, meaning that N is equal in each row; (2) AUC-ROC values are computed using original, non-smoothed ROC curves.", footnote_as_chunk = TRUE, threeparttable = TRUE)
```

```{r}
table2_full <- read_csv("data/table2-for-appendix.csv")

table2_full %>%
  select(horizon, column, N_test, cases) %>%
  pivot_wider(names_from = column, values_from = N_test) %>%
  arrange(horizon, desc(cases)) %>%
  rename(Cases = cases, Horizon = horizon) %>%
  select(Cases, Horizon, everything()) %>%
  arrange(desc(Cases), Horizon) %>%
  knitr::kable("latex", booktabs = TRUE, digits = 0,
               caption = "Number of valid test predictions for each cell in BS Table 2") %>%
  column_spec(3:7, width = "2cm") %>%
  collapse_rows(1:2, row_group_label_position = "stack")

table2_full %>%
  pivot_longer(auc_roc:auc_roc_smoothed, names_to = "Smoothed ROC",
               values_to = "auc_roc") %>%
  mutate(`Smoothed ROC` = case_when(
    `Smoothed ROC`=="auc_roc" ~ "No",
    TRUE ~ "Yes")
  ) %>%
  select(horizon, column, cases, `Smoothed ROC`, auc_roc) %>%
  pivot_wider(names_from = column, values_from = auc_roc) %>%
  rename(Cases = cases, Horizon = horizon) %>%
  select(Cases, Horizon, `Smoothed ROC`, everything()) %>%
  arrange(desc(Cases), Horizon, desc(`Smoothed ROC`)) %>%
  knitr::kable("latex", booktabs = TRUE, digits = 2,
               caption = "Replication of BS Table 2 with smoothed/original ROC and with original varying N cases or adjusting for common case set with constant N") %>%
  column_spec(3, width = "1.5cm") %>%
  column_spec(4:8, width = "2cm") %>%
  collapse_rows(1:3, row_group_label_position = "stack")

```
